---
title: "first"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Harvest Data
Data set for Commercial Harvest from 1997 to 2019 of Iowa Lakes
First we will correct the data set by correcting the date and year columns as well as removing blank cells and renaming cells

```{r}
harvest.dat <- read.csv("harvestdata.csv", TRUE)
harvest.dat$Date <- as.Date(harvest.dat$Date, "m%/d%/Y%")
harvest.dat$Year <- as.factor(harvest.dat$Year)

colnames(harvest.dat)[1] <- c("Lake")

harvest.dat<-droplevels(subset(harvest.dat, Lake != ""))
harvest.dat$Lake[harvest.dat$Lake == "center lake"] <- "Center Lake"
harvest.dat$Lake[harvest.dat$Lake == "Silver Lake (Dickinson)"] <- "Silver Lake"

levels(harvest.dat$Lake)
```

```{r}
center.sub <- subset(harvest.dat, harvest.dat$Lake == "Center Lake")
fiveisland.sub <- subset(harvest.dat, harvest.dat$Lake == "Five Island Lake")
silver.sub <- subset(harvest.dat, harvest.dat$Lake == "Silver Lake")
high.sub <- subset(harvest.dat, harvest.dat$Lake == "High Lake")
ingham.sub <- subset(harvest.dat, harvest.dat$Lake == "Ingham Lake")
lostisland.sub <- subset(harvest.dat, harvest.dat$Lake == "Lost Island Lake")

```

Here are summaries for each lake

```{r}
###Summaries
summary(center.sub)
summary(fiveisland.sub)
summary(high.sub)
summary(silver.sub)
summary(high.sub)
summary(ingham.sub)
summary(lostisland.sub)

```

plotting data for center lake
```{r Center Lake Plots}
plot(center.sub$Date, center.sub$Pounds)
```



# Water Quality Data

Same thing as harvest data but with water quality

```{r}
water.dat <- read.csv("waterdata.csv", TRUE)

#removing the not-needed columns & renaming the needed columns

water.dat <- water.dat[,-c(1,5,8,9,13,20,21)]
colnames(water.dat) <- c("Lake","Date","Year", "Chlorophyll a", "TSS", "TN", "TP", "TP Reactive", "Depth", "Secchi Depth", "Averge Temp", "Average pH", "DO Saturation", "DO", "Turbidity", "Phyotplankton", "Zoops")

#Renaming Each Lake

water.dat$Date <- as.Date(water.dat$Date, "%m/%d/%Y")
water.dat<-droplevels(subset(water.dat, Lake != ""))
water.dat$Lake <- as.character(water.dat$Lake)
water.dat$Lake[water.dat$Lake == "(019) Center Lake (Dickinson ) "] <- "Center Lake"
water.dat$Lake[water.dat$Lake == "(036) Five Island Lake (Palo Alto ) "] <- "Five Island Lake"
water.dat$Lake[water.dat$Lake == "(048) Ingham Lake (Emmet ) "] <- "Ingham Lake"
water.dat$Lake[water.dat$Lake == "(077) Lost Island Lake (Palo Alto ) "] <- "Lost Island Lake"
water.dat$Lake[water.dat$Lake == "(105) Silver Lake (Dickinson) (Dickinson ) "] <- "Silver Lake"
water.dat$Lake <- as.factor(water.dat$Lake)
levels(water.dat$Lake)

```

Subsetting each Lake in the Water Quality data set

```{r}

center.wq <- subset(water.dat, water.dat$Lake == "Center Lake")
fiveisland.wq <- subset(water.dat, water.dat$Lake == "Five Island Lake")
ingham.wq <- subset(water.dat, water.dat$Lake == "Ingham Lake")
lostisland.wq <- subset(water.dat, water.dat$Lake == "Lost Island Lake")
silver.wq <- subset(water.dat, water.dat$Lake == "Silver Lake")

```

```{r}
plot(fiveisland.wq$Date, fiveisland.wq$`Chlorophyll a`)
plot(center.wq$Date, center.wq$Zoops)

```