---
title: "Commercial Fish Harvest ~ Water Quality"
subtitle: "Shallow, natural lakes of northwest Iowa"
output: pdf_document
editor_options: 
  chunk_output_type: console
---


```{r,include=FALSE}
# This chunk clears your environment, if you're like me and it gets messy over there.
rm(list=ls())
```

```{r, include = F}
# This function checks and pre-loads packages for use
packages<-function(x, repos="http://cran.r-project.org", ...){
  x<-as.character(match.call()[[2]])
      if (!require(x,character.only=TRUE)){
          install.packages(pkgs=x, repos=repos, ...)
        require(x,character.only=TRUE)
}
}
# install and attach packages
packages(knitr)
packages(ggplot2)
packages(gridExtra)
packages(reshape2)
packages(tidyr)
packages(rcompanion)
packages(scales)
packages(GGally)

# set figure margins for knitting
knitr::opts_chunk$set(fig.width=8, fig.height=12)

```


# Overall Objective

Our aim is to quantify relationships between commercial harvest of "rough fish" (common carp and bigmouth buffalo) against a suite of water quality variables. The term _water quality_ itself can be subjective, but here we have identified multiple potential response variables that are often associated with "clean" or "impaired" water body designations. They fall into three broad categories: 

- __Nutrients:__ Total Phosphorous ($\mu g/L$); Total Nitrogen ($mg/L$)

- __Physical:__ Secchi depth ($m, meters$); Total Suspended Solids ($\mu g/L$); Dissolved Oxygen ($mg/L$)

- __Biological:__ Total Zooplankton Biomass; Total Phytoplankton Biomass; Phytoplankton Biomass of _Cyanophyta_; Phytoplankton Biomass of _Chlorophyta_

# Study Area

Marty's dissertation research is focused on shallow, natural lakes of Iowa. Commercial Harvest of bigmouth buffalo occurs in some 20 lakes and rivers in Iowa, and over half of them fall into the shallow natural lake category. Since 1) Common carp are known to disturb ecosystems and force shallow lakes into a turbid, algal-dominated state and 2) Bigmouth Buffalo may regulate zooplankton that graze on algae, the Iowa DNR is interested to see if managing these populations of fish can restore elusive water quality in Iowa and shift these lakes into a clear-water macrophyte and sportfish dominated lake systems.

Commercial harvest of these species is complex and variable, with weather conditions, equipment, staffing, invasive species control, permitting, supply chain management, and market fluctuations affecting how much effort commercial anglers can put into catching fish, and where they focus that effort. Fishing is restricted to times of year when water temperatures are below 70$^\circ F$, or basically between Memorial Day and and October-ish. That being said, _there might be some overlap between harvest dates and ambient lake monitoring samples (described below)_.

__One final note on the study area:__ We can incorporate other lakes that fit our selection criteria but do not have commercial harvest of rough fish species. Not that they are a perfect control, but it could help explain any year-to-year variation in response variables across our study region.


# Introducing the Data

## Harvest Data

We do not have abundance data for carp and buffalo for more than a couple of years in a subset of lakes. However, based on available data from the Spirit Lake District Fisheries Office, we have total pounds removed at multiple time points spanning 25 years in about a dozen lakes. We have identified a subset of these lakes based on similarity in geographic location, lake morphometry, fish communities, and regular harvest periods. Data is provided in total pounds removed, however due to differences in lake size and capacity to hold fish, we will standardize data by pounds removed per acre (and also kilograms removed per hectare, for publication). The lakes are summarized here:

Lake                    | Center Lake   | Five Island Lake | High Lake     | Ingham Lake    | Lost Island Lake     | Silver Lake (Dickinson County)     
----------------------- | ------------- | ---------------- | ------------- | -------------- | -------------------- | ---------------------------------
Area (acres)            | 220           | 964              | 467           | 370            | 1,162                | 1,032
Area ($ft^2$)           | 9,583,200     | 41,991,840       | 20,342,520    | 16,117,200     | 50,616,720           | 44,953,920
Watershed Area (acres)  | 731           | 8,689            |               | 320            | 4,541                | 17,025
Wathershed:Lake Ratio   | 3.3:1         | 9.2:1            |               | 0.9:1          | 3.9:1                | 16.5:1
Mean Depth ($ft$)       | 9.5           | 5                |               | 6.2            | 10.2                 | 6.7
Max Depth ($ft$)        | 15.5          | 20               | 8             | 12             | 14                   | 9.8
Row Crop Usage (acres)  |               | 6,565            |               | 135            |                      | 14,521
Percent Row Crop        |               | 75.6             |               | 42.2           |                      | 85.3
Shoreline ($ft$)        | 13,200        | 74,976           |               | 22,400         | 38,544               | 50,730
Shoreline Complexity    | 1.20          | 3.26             |               | 1.57           | 1.53                 | 2.13

- _Note that High Lake is considered a wetland, and not all the same TMDL and shoreline data are available for that lake._
- _For publication, these will need to be converted to metric units._

## Water Quality Data

Water quality data (summarized above) is available from ISU and Iowa DNR through an online data base and the Limnology Laboratory, headed at ISU by Dr. Grace Wilkinson, EEOB. This is part of the Ambient Lake Monitoring Program that collects a suite of variables at 132 lakes in Iowa that are designated as "Principal Recreation Lakes" and data is available from 2000 to 2015, except in 2008. Further data (2016-2019) may also be available for inclusion toward the end of this project's timeframe. These lakes are visited 2-3 (almost always 3) times per year from spring to fall and in-situ measurements are taken as well as samples removed for laboratory analysis of biological and chemical components. 

Lakes fluctuate over time in cycles of stratification, turnover, zooplankton and algal blooms, etc. so these snapshots in time can be problematic. Still, we have identified a suite of variables that we expect to respond to the biomanipulation (harvest) of carp and buffalo. Also, of course, these variables are common metrics used to designate water bodies as "degraded" or "healthy" and could provide insight into food web processes in the lake. The variables are summarized here:


Variable                | Form of Variable   | Units            | Hypothesized Response (short-term) 
----------------------- | ------------------ | ---------------- | -------------------------------------------------------------------
Secchi Depth            | Mean, Maximum      | meters           | Increase from carp removal, no difference from buffalo
Total Suspened Solids   | Mean               | $\mu g/L$        | Decrease from carp removal, no difference from buffalo 
Dissolved Oxygen        | Mean               | $mg/L$           | Decrease if large buffalo removal caused enough increase in zooplankton to reduce O2 production 
                        |                    |                  |
Total Phosphorous       | Mean, Maximum      | $\mu g/L$        | Decrease with carp removal (resuspension)
Total Nitrogen          | Mean, Maximum      | $mg/L$           | Decrease with carp removal (resuspension)
                        |                    |                  |
Zooplankton Biomass     | Mean               | $\mu g/L$        | Increase with buffalo removal (trophic release)
Chlorophyll _a_         | Mean, Maximum      | $\mu g/L$        | Decrease with carp and buffalo removal
Phytoplankton Biomass   | Mean               | $mg/L$           | Decrease with both carp and buffalo removal (bottom up and top down pathways)
_Cyanophyta_            | Mean               | $mg/L$           | Decrease with carp and buffalo removal (nutrient shift and zoop shift)
_Chlorophyta_           | Mean               | $mg/L$           | Decrease with both carp and buffalo removal

__Trophic State Index__ | Calculated from Chl _a_ | Index Value | Decrease with both carp and buffalo removal; but not out of hyper-eurtrophic state.

- _A Trophic State Index (TSI) can also be calculated for Secchi Depth, and Total Phosphorus._ We will choose to pick chlorophyll _a_ as the response variable, because the developer of the index (Carlson 1977) said in 1983 that averaging the three values was not appropriate, and chlorophyll _a_ was the best measurement to use out of the three. We do not expect the TSI to decrease out of the eutrophic range, but a directional shift in TSI may be indicative of the desired change in water quality from biomanipulation of rough fish.


## Data Analysis

### Loading in Harvest Data

- Data set for Commercial Harvest from 1997 to 2019 of Iowa Lakes
- Plot the distribution of harvest in total pounds removed (left) and pounds per acre removed (right). The number of bars across the box plots indicate the number of harvest dates in that year due to "stacking" within a year. The colors represent the different species.

```{r, include = F}
harvest.dat <- read.csv("harvestdata.csv", TRUE)
head(harvest.dat)
harvest.dat$Date <- as.Date(harvest.dat$Date, "%m/%d/%Y")
harvest.dat$Year <- as.factor(harvest.dat$Year)
harvest.dat$Month<-as.factor(harvest.dat$Month)
harvest.dat$Modified.Year<-as.factor(harvest.dat$Modified.Year)
colnames(harvest.dat)[1] <- c("Lake")

harvest.dat<-droplevels(subset(harvest.dat, Lake != ""))
harvest.dat$Lake<-as.character(harvest.dat$Lake)
harvest.dat$Lake[harvest.dat$Lake == "center lake"] <- "Center Lake"
harvest.dat$Lake[harvest.dat$Lake == "Silver Lake (Dickinson)"] <- "Silver Lake"
harvest.dat$Lake<-as.factor(harvest.dat$Lake)
levels(harvest.dat$Lake) # 6 lakes all good


# setting acres
harvest.dat$Acres[harvest.dat$Lake == "Center Lake"] <- 220
harvest.dat$Acres[harvest.dat$Lake == "Five Island Lake"] <- 964
harvest.dat$Acres[harvest.dat$Lake == "High Lake"] <- 467
harvest.dat$Acres[harvest.dat$Lake == "Ingham Lake"] <- 370
harvest.dat$Acres[harvest.dat$Lake == "Lost Island Lake"] <- 1162
harvest.dat$Acres[harvest.dat$Lake == "Silver Lake"] <- 1032


str(harvest.dat)

# not sure about the last two columns before acres::
harvest.dat<-harvest.dat[,-c(10,11)]


# There's a period ( -> . <- ) in one cell in pounds column, making it a factor
harvest.dat$Pounds[harvest.dat$Pounds == "."] <- 0
harvest.dat$Pounds<-as.numeric(as.character(harvest.dat$Pounds))
summary(harvest.dat$Pounds)


###############################################
#
#    Predictor variable standardized by lake: pounds per acre; kg/ha
#

harvest.dat$Std.Removal.lbs<-harvest.dat$Pounds/harvest.dat$Acres
harvest.dat$Std.Removal.kg<-harvest.dat$Std.Removal.lbs/1.12085 

# Delete extraneous info: angler
harvest.dat<-harvest.dat[,-c(2,3)]

# Subsetting to only data from June 1999 onward (This is the 1999-2000 harvest season, before start of ALM water quality data)
harvest.dat<-droplevels(subset(harvest.dat, Date >= "1999-06-01"))
str(harvest.dat)
```

```{r, include = F}
# subset frames if needed later
center.sub <- droplevels(subset(harvest.dat, harvest.dat$Lake == "Center Lake"))
fiveisland.sub <- droplevels(subset(harvest.dat, harvest.dat$Lake == "Five Island Lake"))
silver.sub <- droplevels(subset(harvest.dat, harvest.dat$Lake == "Silver Lake"))
high.sub <- droplevels(subset(harvest.dat, harvest.dat$Lake == "High Lake"))
ingham.sub <- droplevels(subset(harvest.dat, harvest.dat$Lake == "Ingham Lake"))
lostisland.sub <- droplevels(subset(harvest.dat, harvest.dat$Lake == "Lost Island Lake"))

```


```{r, include = F}
###Summaries
summary(center.sub) 
summary(fiveisland.sub)
summary(high.sub)
summary(silver.sub)
summary(high.sub)
summary(ingham.sub)
summary(lostisland.sub)

```

### Harvest through time
```{r, include = F}

a <- ggplot(harvest.dat, aes(x=Year,y=Pounds, fill = Species))+
            geom_bar(stat= "identity", colour = "black")+
                labs(title = "Harvest in Pounds through Time",
                 x="Date",
                 y="Pounds Removed") +
            facet_grid(Lake~.)+
            scale_y_continuous(labels=comma)+
            theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = "bottom",
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))




b <- ggplot(harvest.dat, aes(x=Year,y= Std.Removal.lbs, fill = Species))+
            geom_bar(stat= "identity", colour = "black")+
                labs(title = "Harvest in Pounds per Acre through Time",
                 x="Date",
                 y="Pounds per Acre Removed") +
            facet_grid(Lake~.)+
            scale_y_continuous(labels=comma)+
            theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = "bottom",
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))

grid.arrange(a,b,ncol=2)
```



## Loading in  Water Quality Data

```{r, include = F}
water.dat <- read.csv("waterdata.csv", TRUE)

#removing the not-needed columns & renaming the needed columns

water.dat <- water.dat[,-c(1,5,8,9,13,20,21)] # M Foli CHeck these two lines for accuracy please!
colnames(water.dat) <- c("Lake","Date","Year", "Chlorophyll a", "TSS", "TN", "TP", "TP Reactive", "Depth", "Secchi Depth", "Averge Temp", "Average pH", "DO Saturation", "DO", "Turbidity", "Phyotplankton", "Zoops", "Chlorophyta", "Cyanophyta")

#Renaming Each Lake

water.dat$Date <- as.Date(water.dat$Date, "%m/%d/%Y")
water.dat<-droplevels(subset(water.dat, Lake != ""))
water.dat$Lake <- as.character(water.dat$Lake)
water.dat$Lake[water.dat$Lake == "(019) Center Lake (Dickinson ) "] <- "Center Lake"
water.dat$Lake[water.dat$Lake == "(036) Five Island Lake (Palo Alto ) "] <- "Five Island Lake"
water.dat$Lake[water.dat$Lake == "(048) Ingham Lake (Emmet ) "] <- "Ingham Lake"
water.dat$Lake[water.dat$Lake == "(077) Lost Island Lake (Palo Alto ) "] <- "Lost Island Lake"
water.dat$Lake[water.dat$Lake == "(105) Silver Lake (Dickinson) (Dickinson ) "] <- "Silver Lake"
water.dat$Lake <- as.factor(water.dat$Lake)
levels(water.dat$Lake)

# Year as factor
water.dat$Year<-as.factor(water.dat$Year)

str(water.dat)
```


```{r, include = F}
# Subsetting each Lake in the Water Quality data set, if needed later
center.wq <- droplevels(subset(water.dat, water.dat$Lake == "Center Lake"))
fiveisland.wq <- droplevels(subset(water.dat, water.dat$Lake == "Five Island Lake"))
ingham.wq <- droplevels(subset(water.dat, water.dat$Lake == "Ingham Lake"))
lostisland.wq <- droplevels(subset(water.dat, water.dat$Lake == "Lost Island Lake"))
silver.wq <- droplevels(subset(water.dat, water.dat$Lake == "Silver Lake"))
high.wq <- droplevels(subset(water.dat, water.dat$Lake == "High Lake"))
```

### Distribution of water quality variables, for each Lake 

```{r, include = F}
#summary(water.dat)
# first plot: TSS, Secchi, DO (so three columns, 6 rows)
water.dat.m.a<-melt(water.dat[,c(1,5,10,14)])

a<-ggplot(water.dat.m.a, aes(x=value))+
      geom_histogram(colour = "black", fill = "white")+
      geom_density(alpha=.2, fill = "#FF6666")+
            facet_grid(Lake~variable, scales = "free")+
            theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = "bottom",
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))
a

# Second plot: TP and TN (two columns, thre rows)
water.dat.m.a<-melt(water.dat[,c(1,6,7)])

b<-ggplot(water.dat.m.a, aes(x=value))+
      geom_histogram(colour = "black", fill = "white")+
      geom_density(alpha=.2, fill = "#FF6666")+
            facet_grid(Lake~variable, scales = "free")+
            theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = "bottom",
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))
b

# Third plot: Zoops, Cholorophyll, Phytoplankton, chlorophyta and cyanophyta
water.dat.m.a<-melt(water.dat[,c(1,17,18,19,16,4)])

c<-ggplot(water.dat.m.a, aes(x=value))+
      geom_histogram(colour = "black", fill = "white")+
      geom_density(alpha=.2, fill = "#FF6666")+
            facet_grid(Lake~variable, scales = "free")+
            theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = "bottom",
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))
c


```

## Update 2/24/2020

After meeting with Audrey and Phil, we have a new format for how to arrange the data. The general model format is:

\[
TSI \approx \beta_0 + L_i + (\beta_c * C_{li} * H_i) + (\beta_b * (1-C_{li}) * H_i)
\]

where i is indexed on year, l is indexed on lake, L is a fixed lake effect, C is the proportion of harvest made up by carp, 1-C is the proportion of harvest made up by buffalo, H is the total harvest. 

Action Items:

- Calculate TSI Chla taken from the June measurements. When a measurement from June is not available, we took the next closest measurement to one full year after the previous year's measurement date. 

- From those TSI values, calculate the chanage in TSI from year a to year b. 

- Apply same change in variable between year a and year b for other WQ variables.

### New data frame

- Lake

- Year*: Year B for water quality stuff (represents the change in Y between year A and B)
- Year*: Total and proportional harvests between years a and b (September A to April B)

- Include the amount (pounds) of carp harvested and buffalo harvested between years A and B

- Create a new variable of the proportion of total harvest (lbs of species / lbs of both species) for carp and buffalo between years A and B

- TSI and other response variables, as the change in Y from year A to year B

------------  So the overall data frame will look like:
Lake| Year* | Pounds Carp | Pounds Buff | Prop. harvested carp | Prop. harvested buff | TSI | Y variables

## Loading in Yearly Change Dataset

```{r, echo = F, warning=F}
yearchange.dat <- read.csv("Changesdata.csv", TRUE) 
colnames(yearchange.dat)[1] <- c("Lake")


#Setting Year as factor
yearchange.dat$Year <- as.factor(yearchange.dat$Year)


#Subsetting each lake for future use
center.yc <- droplevels(subset(yearchange.dat, yearchange.dat$Lake == "Center Lake"))
fiveisland.yc <- droplevels(subset(yearchange.dat, yearchange.dat$Lake == "Five Island Lake"))
ingham.yc <- droplevels(subset(yearchange.dat, yearchange.dat$Lake == "Ingham Lake"))
lostisland.yc <- droplevels(subset(yearchange.dat, yearchange.dat$Lake == "Lost Island Lake"))
silver.yc <- droplevels(subset(yearchange.dat, yearchange.dat$Lake == "Silver Lake"))
high.yc <- droplevels(subset(yearchange.dat, yearchange.dat$Lake == "High Lake"))

```

### Distribution of response variables: TSS, Secchi Depth, Dissolved Oxygen

These represent the count(frequency) for each response variable along it's range. geom_density from ggplot2 is a smoothed version of the histogram, useful for continuous random variables.

```{r, echo = F, warning=F}
#Plotting TSS, Secchi, DO
head(yearchange.dat)
change.dat.m.a<-melt(yearchange.dat[,c(1,10,13,14)]) # TSS col 10, Secchi col 13, 

a<-ggplot(change.dat.m.a, aes(x=value))+
      geom_histogram(colour = "black", fill = "white")+
      geom_density(alpha=.2, fill = "#FF6666")+
            facet_grid(Lake~variable, scales = "free")+
            theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = "bottom",
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))
a

#Plotting TN and TP

change.dat.m.b<-melt(yearchange.dat[,c(1,11,12)]) # TN and TP

b<-ggplot(change.dat.m.b, aes(x=value))+
      geom_histogram(colour = "black", fill = "white")+
      geom_density(alpha=.2, fill = "#FF6666")+
            facet_grid(Lake~variable, scales = "free")+
            theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = "bottom",
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))
b

#Plotting TSI, Cholorophyll, Phytoplankton, Zoops, chlorophyta, and cyanophyta

change.dat.m.c<-melt(yearchange.dat[,c(1,8,9,16,17,18,19)])

c<-ggplot(change.dat.m.c, aes(x=value))+
      geom_histogram(colour = "black", fill = "white")+
      geom_density(alpha=.2, fill = "#FF6666")+
            facet_grid(Lake~variable, scales = "free")+
            theme(panel.grid.major = element_blank(), 
                              panel.grid.minor = element_blank(),
                              panel.background = element_rect(fill="white",colour="grey50"), 
                              axis.line = element_line(colour = "black"),
                              legend.position = "bottom",
                              legend.title = element_blank(),
                              legend.background = element_rect(colour="grey20"))
c
```


#### We can drop High lake if the data is problematic


### pairs plot from GGally

```{r,}
# ggpairs(yearchange.dat[,8:19])
# too many!

ggpairs(yearchange.dat[,8:19])

```



