---
title: "first"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Harvest Data
Data set for Commercial Harvest from 1997 to 2019 of Iowa Lakes
First we will correct the data set by correcting the date and year columns as well as removing blank cells and renaming cells

```{r}
harvest.dat <- read.csv("harvestdata.csv", TRUE)
harvest.dat$Date <- as.Date(harvest.dat$Date, "m%/d%/Y%")
harvest.dat$Year <- as.factor(harvest.dat$Year)

colnames(harvest.dat)[1] <- c("Lake")

harvest.dat<-droplevels(subset(harvest.dat, Lake != ""))
harvest.dat$Lake[harvest.dat$Lake == "center lake"] <- "Center Lake"
harvest.dat$Lake[harvest.dat$Lake == "Silver Lake (Dickinson)"] <- "Silver Lake"

levels(harvest.dat$Lake)
```

```{r}
center.sub <- subset(harvest.dat, harvest.dat$Lake == "Center Lake")
fiveisland.sub <- subset(harvest.dat, harvest.dat$Lake == "Five Island Lake")
silver.sub <- subset(harvest.dat, harvest.dat$Lake == "Silver Lake")
high.sub <- subset(harvest.dat, harvest.dat$Lake == "High Lake")
ingham.sub <- subset(harvest.dat, harvest.dat$Lake == "Ingham Lake")
lostisland.sub <- subset(harvest.dat, harvest.dat$Lake == "Lost Island Lake")

```

Here are summaries for each lake

```{r}
###Summaries
summary(center.sub)
summary(fiveisland.sub)
summary(high.sub)
summary(silver.sub)
summary(high.sub)
summary(ingham.sub)
summary(lostisland.sub)

```

plotting data for center lake
```{r Center Lake Plots}
plot(center.sub$Date, center.sub$Pounds)
```



# Water Quality Data

Same thing as harvest data but with water quality

```{r}
water.dat <- read.csv("waterdata.csv", TRUE)

#removing the not-needed columns & renaming the needed columns

water.dat <- water.dat[,-c(1,5,8,9,13,20,21)]
colnames(water.dat) <- c("Lake","Date","Year", "Chlorophyll a", "TSS", "TN", "TP", "TP Reactive", "Depth", "Secchi Depth", "Averge Temp", "Average pH", "DO Saturation", "DO", "Turbidity", "Phyotplankton", "Zoops", "Chlorophyta", "Cyanophyta")

#Renaming Each Lake

water.dat$Date <- as.Date(water.dat$Date, "%m/%d/%Y")
water.dat<-droplevels(subset(water.dat, Lake != ""))
water.dat$Lake <- as.character(water.dat$Lake)
water.dat$Lake[water.dat$Lake == "(019) Center Lake (Dickinson ) "] <- "Center Lake"
water.dat$Lake[water.dat$Lake == "(036) Five Island Lake (Palo Alto ) "] <- "Five Island Lake"
water.dat$Lake[water.dat$Lake == "(048) Ingham Lake (Emmet ) "] <- "Ingham Lake"
water.dat$Lake[water.dat$Lake == "(077) Lost Island Lake (Palo Alto ) "] <- "Lost Island Lake"
water.dat$Lake[water.dat$Lake == "(105) Silver Lake (Dickinson) (Dickinson ) "] <- "Silver Lake"
water.dat$Lake <- as.factor(water.dat$Lake)
levels(water.dat$Lake)

```

Subsetting each Lake in the Water Quality data set

```{r}

center.wq <- subset(water.dat, water.dat$Lake == "Center Lake")
fiveisland.wq <- subset(water.dat, water.dat$Lake == "Five Island Lake")
ingham.wq <- subset(water.dat, water.dat$Lake == "Ingham Lake")
lostisland.wq <- subset(water.dat, water.dat$Lake == "Lost Island Lake")
silver.wq <- subset(water.dat, water.dat$Lake == "Silver Lake")
high.wq <- subset(water.dat, water.dat$Lake == "High Lake")
```

Boxplots of all water quality variables for each Lake 

```{r}

#Nutrient bois

boxplot(water.dat$`Chlorophyll a`~water.dat$Lake, ylab='Chlorphyll a (ug/L)', xlab='Lake', main= 'Chlorphyll a Concentrations of Each Lake')
boxplot(water.dat$TN~water.dat$Lake, ylab='Total Nitrogen (mg/L)', xlab='Lake', main= 'Total Nitrogen Concentrations of Each Lake')
boxplot(water.dat$TP~water.dat$Lake, ylab='Total Phosphorous (ug/L)', xlab='Lake', main= 'Total Phosphorous Concentrations of Each Lake')

#Water Clairty

boxplot(water.dat$`Secchi Depth`~water.dat$Lake, ylab= 'Secchi Depth (m)', xlab= 'Lake', main = 'Secchi Depth of Each Lake')
boxplot(water.dat$TSS~water.dat$Lake, ylab= 'Total Suspended Solids (mg/L)', xlab= 'Lake', main = 'Total Suspended Solids of Each Lake')
boxplot(water.dat$DO~water.dat$Lake, ylab= 'Dissolved Oxygen (mg/L)', xlab= 'Lake', main = 'Dissolved Oxygen of Each Lake')
boxplot(water.dat$Turbidity~water.dat$Lake, ylab= 'Turbidity (NTU)', xlab= 'Lake', main = 'Turbidity of Each Lake')


#swimmy bois
boxplot(water.dat$Phyotplankton~water.dat$Lake, ylab= 'Phytoplankton (mg/L)', xlab= 'Lake', main = 'Phytoplankton Densities of Each Lake')
boxplot(water.dat$Zoops~water.dat$Lake, ylab= 'Zooplankton (mg/L)', xlab= 'Lake', main = 'Zooplankton Densities of Each Lake')
boxplot(water.dat$Chlorophyta~water.dat$Lake, ylab= 'Chlorophyta (mg/L)', xlab= 'Lake', main= 'Chlorophyta Densities of Each Lake')
boxplot(water.dat$Cyanophyta~water.dat$Lake, ylab= 'Cyanophyta (mg/L)', xlab= 'Lake', main= 'Cyanophyta Densities of Each Lake')

```